---
title: 'Permanova: Real Data Analysis'
author: "Paul Harmon"
date: "6/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("OutlierCompFunctions.R")
library(ggplot2)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(janitor)
library(magrittr)
library(knitr)
library(Rtsne)
library(robustHD)
library(GGally)
library(stringr)
library(plotly)
library(R.utils)
library(vegan)
library(mnormt)
```

# Introduction

This document overviews the real data analysis portion of the t-SNE/MDS/Permanova paper. We are currently looking at two specific datasets: Palmer Penguins and Carnegie Classifications. 


# Carnegie

Looks at the 2015 Carnegie Classification data - without doing much, to see how the data are changed. 

## Raw Data

```{r, cache = TRUE}
CC2015 <- read.csv("https://raw.githubusercontent.com/paulharmongj/Carnegie_SEM/master/data/CC2015data.csv")

CC2015_filter <- CC2015 %>% filter(BASIC2015 %in% c(15,16,17)) %>% select(PDNFRSTAFF, S.ER.D, NONS.ER.D, FACNUM, HUM_RSD, STEM_RSD, OTHER_RSD, SOCSC_RSD)
perplexity = 30

out <- MultiPermanova(CC2015_filter)
```


```{r}
#generate some plots of the output
df3 = tibble(Holdout = 1:length(out$cmdlist), PValues = sapply(out$cmdlist, pullPval), Fstat = sapply(out$cmdlist, pullFstat))
df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)

  
  
  #Visualize the results and store in a list
  sigplot2 <- ggplot(df3, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: P-Values From Adonis")
  sigplot2
  
  
  cc <- CC2015 %>% filter(BASIC2015 %in% c(15,16,17)) 
  cc$NAME[which(df3$PValues < 0.05)]
  

```



## Scaled

This version looks at the same input variables but considers the set of features that are scaled in the per-capita version. Due to high degrees of correlation between the raw inputs and some of their per-capita counterparts, I only included the per-capita versions for STEM expenditures, Non-STEM research expenditures, and tenurable research faculty count. 

```{r, cache = TRUE
}
CC2015_filter_PC <- CC2015_filter %>% mutate(S.ER.DPC = S.ER.D/FACNUM, NONS.ER.DPC = NONS.ER.D/FACNUM, STAFFPC = PDNFRSTAFF/FACNUM)

out2 <- MultiPermanova(CC2015_filter_PC)

```


## "Extreme" points added in

The goal here is to add a couple "extreme" points and see how this method handles the original "influential" points. 

```{r, eval = FALSE}

```



# Palmer 